---
title: "Myelin Development Sensitivity Analysis: Structural Data Quality"
author: "Valerie Jill Sydnor"
output: 
  rmdformats::material:
    highlight: monochrome
    lightbox: true
    gallery: true
    css: /Volumes/Hera/Projects/corticalmyelin_development/code/corticalmyelin_maturation/results/custom_css/myelindev.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyr)
library(plyr)
library(ggplot2)
library(ggseg)
library(ggsegGlasser)
library(purrr)
library(matrixStats)
library(scales)
library(cifti)
library(ggpubr)
library(paletteer)
library(dplyr)
library(ggnewscale)
library(EnvStats)
library(ggcorrplot)
extrafont::loadfonts()
source("/Volumes/Hera/Projects/corticalmyelin_development/code/corticalmyelin_maturation/gam_models/gam_functions.R")
```

# Read in Data

Glasser regions and assignments

```{r}
#Glasser region and label names for the frontal lobe
glasser.regions <- read.csv("/Volumes/Hera/Projects/corticalmyelin_development/Maps/HCPMMP_glasseratlas/glasser360_regionlist.csv")
glasser.frontal <- read.csv("/Volumes/Hera/Projects/corticalmyelin_development/Maps/HCPMMP_glasseratlas/glasser360_regionlist_frontallobe.csv")
glasser.snr.exclude <- read.csv("/Volumes/Hera/Projects/corticalmyelin_development/Maps/SNR/glasser_SNR_exclusion.csv")
```

```{r}
glasser.plotting <- glasser.regions
glasser.plotting$cortex <- "cortex"
glasser.plotting$cortex[glasser.plotting$orig_parcelname %in% glasser.snr.exclude$orig_parcelname] <- "exclude"
glasser.plotting <- glasser.plotting %>% select(label, cortex)
glasser.plotting <- rbind(glasser.plotting, data.frame(label = "rh_???", cortex = "exclude"))
glasser.plotting <- rbind(glasser.plotting, data.frame(label = "lh_???", cortex = "exclude"))
glasser.plotting <- glasser.plotting %>% filter(label != "lh_L_10pp")
```

Depths

```{r}
depth.list <- c("superficial", "middle", "deep")
ordered_depths <- c("deep", "middle", "superficial")
depth_colorbar <-  c("#004A38FF", "#6992CC", "#A29DC4")
depth_colorbar_reverse <-  c("#A29DC4", "#6992CC", "#004A38FF")
```

S-A Axis

```{r, warning=F}
#S-A axis or saxis as nicknamed by Dan Margulies on a trip to gingerworld
S.A.axis.cifti <- read_cifti("/Volumes/Hera/Projects/corticalmyelin_development/Maps/S-A_ArchetypalAxis/FSLRVertex/SensorimotorAssociation_Axis_parcellated/SensorimotorAssociation.Axis.Glasser360.pscalar.nii") #S-A_ArchetypalAxis repo
S.A.axis <- data.frame(SA.axis = rank(S.A.axis.cifti$data), orig_parcelname = names(S.A.axis.cifti$Parcel))
S.A.axis <- merge(S.A.axis, glasser.regions, by = "orig_parcelname")
```

Final participant list

```{r}
participants <- read.csv("/Volumes/Hera/Projects/corticalmyelin_development/sample_info/7T_MP2RAGE_finalsample_demographics.csv")
participants <- participants %>% mutate(subses = sprintf("%s_%s", subject_id, session_id))
```

GAM outputs: developmental effects

```{r}
setwd("/Volumes/Hera/Projects/corticalmyelin_development/gam_outputs/sensitivity_analyses/") #output from /gam_models/fit_sensitivityGAMs_glasserregions_bydepth.R

files <- list.files(getwd(), pattern = "age_euler.RDS") 

#read in files and assign to variables
for(i in 1:length(files)){
  
  Rfilename <- gsub(".RDS", "", files[i])
  
  x <- readRDS(files[i]) 
  assign(Rfilename, x) 
  }
```

```{r}
gam.results.alldepths <- list(superficial_gamstatistics_age_euler, middle_gamstatistics_age_euler, deep_gamstatistics_age_euler)
names(gam.results.alldepths) <- depth.list
```

```{r}
#Extract the 4 types of results and format
gam.statistics.alldepths <- lapply(gam.results.alldepths, '[[', "gam.statistics.df" )
gam.fittedvalues.alldepths <-  lapply(gam.results.alldepths, '[[', "gam.fittedvalues.df" )
gam.smoothestimates.alldepths <- lapply(gam.results.alldepths, '[[', "gam.smoothestimates.df" )
gam.derivatives.alldepths <- lapply(gam.results.alldepths, '[[', "gam.derivatives.df" )

format_depth_dfs <- function(depth){
  depth <- depth[depth$orig_parcelname %in% glasser.frontal$orig_parcelname,] #get results for frontal lobe only
  depth <-  depth[!(depth$orig_parcelname %in% glasser.snr.exclude$orig_parcelname),] #exclude low SNR parcels
}

gam.statistics.alldepths <- lapply(gam.statistics.alldepths, function(depth){
  format_depth_dfs(depth)
})
gam.fittedvalues.alldepths <- lapply(gam.fittedvalues.alldepths, function(depth){
  format_depth_dfs(depth)
})
gam.smoothestimates.alldepths <- lapply(gam.smoothestimates.alldepths, function(depth){
  format_depth_dfs(depth)
})
gam.derivatives.alldepths <- lapply(gam.derivatives.alldepths, function(depth){
  format_depth_dfs(depth)
})
```

# Myelin Develops Heterochronously Across Superficial, Middle, and Deep Cortex

## Number of significant regional effects at each depth

```{r}
#Function to calculate the number of significant age effects post-FDR
ageeffect.results <- function(depth){
  ageeffects <- depth
  n_roi <- nrow(ageeffects)
 
  #significant smooth terms
  aageeffects <- ageeffects %>% mutate(significant = p.adjust(ageeffects$GAM.smooth.pvalue, method = c("fdr")) < 0.05)
  sig_age <- aageeffects %>% filter(significant == TRUE) %>% nrow()
  sig_age_percent <- round((sig_age/n_roi)*100, 2)
                         
  sprintf("%s percent of regions show a positive significant developmental change in R1", sig_age_percent)
}
```

```{r}
lapply(gam.statistics.alldepths, function(depth){
  ageeffect.results(depth)
})
```

## Regional derivatives at each depth

```{r}
gam.derivatives.alldepths.long <- do.call(rbind, gam.derivatives.alldepths)
gam.derivatives.alldepths.long <- gam.derivatives.alldepths.long %>% mutate(depth = factor(str_remove(row.names(gam.derivatives.alldepths.long), "\\..*")))
gam.derivatives.alldepths.long <- merge(gam.derivatives.alldepths.long, glasser.regions, sort = F)

#Calculate the average first derivative in each region at each depth
regional.rate.alldepths <- gam.derivatives.alldepths.long %>% group_by(label, depth) %>% do(rate = mean(.$derivative)) %>% unnest(cols = "rate")

#Calculate the average across-region first derivative at each depth
frontallobe.rate.alldepths <- regional.rate.alldepths %>% group_by(depth) %>% do(mean.rate = mean(.$rate)) %>% unnest(cols = "mean.rate")
frontallobe.rate.alldepths$depth <- c("D", "M", "S")
```

**Depth-wise derivative plot**

```{r}
frontallobe.rate.alldepths %>% 
  ggplot(aes(x = mean.rate, y = depth, fill = depth)) +
  geom_point(shape = 23, size = 3, color = "white") +
  theme_classic() +
  scale_fill_manual(values = depth_colorbar) +
  xlab("\nRate (derivative)") +
  ylab("Cortical compartment\n") +
  scale_x_continuous(breaks = c(0.001, 0.002), limits = c(0.001, 0.0021)) +
  theme(
  legend.position = "none",
  axis.text = element_text(size = 6, family = "Arial", color = c("black")),
  axis.title.x = element_text(size = 6, family ="Arial", color = c("black")),
  axis.title.y = element_text(size = 6, family ="Arial", color = c("black")),
  axis.line = element_line(linewidth = .2), 
  axis.ticks = element_line(linewidth = .2)) 

ggsave(filename = "/Users/valeriesydnor/Documents/ResearchProjects/CorticalMyelin_PlasticityDevelopment/Publications/Publication_Images/Figure6/Euler_depthplot_derivative.pdf", device = "pdf", dpi = 500, width = 1.1, height = 1.3)
```

**Depth-wise derivative brain plots**

```{r, message = F}
for(this.depth in c("superficial", "deep")){
  
  regional.rate.depth <- regional.rate.alldepths %>% filter(depth == this.depth) %>% filter(label != "lh_L_10pp")
  
  derivative.plot <- ggplot() + 
  geom_brain(data = regional.rate.depth, atlas = glasser, mapping = aes(fill = rate), colour=I("gray50"), size=I(.08)) + theme_classic() + theme(legend.position = "none") + 
  paletteer::scale_fill_paletteer_c("grDevices::PuBuGn", direction = -1, limits = c(0.0005, 0.0022), oob = squish, na.value = "white") +  
  theme(legend.position = "none", axis.text = element_blank(), axis.line = element_blank(), axis.ticks = element_blank()) +
  new_scale_fill() + 
  geom_brain(data = glasser.plotting, atlas = glasser, mapping = aes(fill = cortex),  colour=I(alpha("gray50", 0)), size=I(0)) +
  scale_fill_manual(values = c(alpha("white", 0), "gray75")) 
  
  print(derivative.plot)
  
ggsave(filename = sprintf("/Users/valeriesydnor/Documents/ResearchProjects/CorticalMyelin_PlasticityDevelopment/Publications/Publication_Images/Figure6/Euler_%s_rate_corticalmap.pdf", this.depth), device = "pdf", dpi = 500, width = 3.15, height = 1.7)
}
```

## Regional age of maturation at each depth

```{r}
gam.statistics.alldepths.long <- do.call(rbind, gam.statistics.alldepths)
regional.maturation.alldepths <- gam.statistics.alldepths.long %>% mutate(depth = factor(str_remove(row.names(gam.statistics.alldepths.long), "\\..*")))
regional.maturation.alldepths <- merge(regional.maturation.alldepths, glasser.regions, sort = F)

#Calculate the average across-region age of maturation at each depth
frontallobe.maturation.alldepths <- regional.maturation.alldepths %>% group_by(depth) %>% do(mean.age = mean(.$smooth.increase.offset, na.rm = T)) %>% unnest(cols = "mean.age")
frontallobe.maturation.alldepths$depth <- c("D", "M", "S")
```

**Depth-wise age of maturation plot**

```{r}
frontallobe.maturation.alldepths %>% 
  ggplot(aes(x = mean.age, y = depth, fill = depth)) +
  geom_point(shape = 23, size = 3, color = "white") +
  theme_classic() +
  scale_fill_manual(values = depth_colorbar) +
  xlab("\nMaturation (years)") +
  ylab("Cortical compartment\n") +
  scale_x_continuous(limits = c(25.7, 30.5), breaks = c(26, 28, 30)) +
  theme(
  legend.position = "none",
  axis.text = element_text(size = 6, family = "Arial", color = c("black")),
  axis.title.x = element_text(size = 6, family ="Arial", color = c("black")),
  axis.title.y = element_text(size = 6, family ="Arial", color = c("black")),
  axis.line = element_line(linewidth = .2), 
  axis.ticks = element_line(linewidth = .2)) 

ggsave(filename = "/Users/valeriesydnor/Documents/ResearchProjects/CorticalMyelin_PlasticityDevelopment/Publications/Publication_Images/Figure6/Euler_depthplot_maturation.pdf", device = "pdf", dpi = 500, width = 1.1, height = 1.3)
```

**Depth-wise age of maturation brain plots**

```{r, message = F}
for(this.depth in c("superficial", "deep")){
  
  regional.mat.depth <- regional.maturation.alldepths %>% filter(depth == this.depth) %>% filter(label != "lh_L_10pp")
  
  derivative.plot <- ggplot() + 
  geom_brain(data = regional.mat.depth, atlas = glasser, mapping = aes(fill = smooth.increase.offset), colour=I("gray50"), size=I(.08)) + theme_classic() + theme(legend.position = "none") + 
  paletteer::scale_fill_paletteer_c("grDevices::PuBuGn", direction = 1, limits = c(26, 34), oob = squish, na.value = "white") +  
  theme(legend.position = "none", axis.text = element_blank(), axis.line = element_blank(), axis.ticks = element_blank()) +
  new_scale_fill() + 
  geom_brain(data = glasser.plotting, atlas = glasser, mapping = aes(fill = cortex),  colour=I(alpha("gray50", 0)), size=I(0)) +
  scale_fill_manual(values = c(alpha("white", 0), "gray75")) 
  
  print(derivative.plot)
  
ggsave(filename = sprintf("/Users/valeriesydnor/Documents/ResearchProjects/CorticalMyelin_PlasticityDevelopment/Publications/Publication_Images/Figure6/Euler_%s_maturation_corticalmap.pdf", this.depth), device = "pdf", dpi = 500, width = 3.15, height = 1.7)
}
```

# Depth-wise Maturational Profiles Reflect Functional Diversity

## Depth-dependent trajectories in exemplar regions

```{r}
gam.smoothestimates.alldepths.long <- do.call(rbind, gam.smoothestimates.alldepths)
gam.smoothestimates.alldepths.long <- gam.smoothestimates.alldepths.long %>% mutate(depth = factor(str_remove(row.names(gam.smoothestimates.alldepths.long), "\\..*")))
```

```{r}
#Function to plot age smooth functions in exemplar regions
plot.depth.smooths <- function(region, y_ticks){

  gam.smoothestimates.alldepths.long$depth <- factor(gam.smoothestimates.alldepths.long$depth, ordered = T, levels = depth.list)
  
  smooth.plot <- gam.smoothestimates.alldepths.long %>% filter(orig_parcelname == region) %>%
  ggplot(., aes(x = age, y = est, group = depth, color = depth)) +
  geom_line(linewidth = .35) +
  theme_classic() +
  xlab("\nAge (years)") +
  ylab("R1 (centered)\n") +
  scale_y_continuous(breaks = y_ticks) +
  scale_x_continuous(breaks = c(10, 20, 30)) +
  scale_color_manual(values = depth_colorbar_reverse) +
  theme(
        legend.position = "none", 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_text(size = 6, family = "Arial", color = c("black")),
        axis.title.x = element_text(size = 7, family ="Arial", color = c("black")),
        axis.title.y = element_text(size = 7, family ="Arial", color = c("black")),
        axis.line = element_line(linewidth = .2), 
        axis.ticks = element_line(linewidth = .2))
  
  return(smooth.plot)
}
```

**Primary Motor (4)**

```{r}
plot.depth.smooths(region = "L_4_ROI", y_ticks = c(-0.04, -0.02, 0, 0.02))
ggsave(filename = "/Users/valeriesydnor/Documents/ResearchProjects/CorticalMyelin_PlasticityDevelopment/Publications/Publication_Images/Figure6/Euler_Area4_depthsmooths.pdf", device = "pdf", dpi = 500, width = 1.5, height = 1.4)
```

**Dorsolateral PFC (46)**

```{r}
plot.depth.smooths(region = "L_46_ROI", y_ticks = c(-0.02, -0.01, 0, 0.01))
ggsave(filename = "/Users/valeriesydnor/Documents/ResearchProjects/CorticalMyelin_PlasticityDevelopment/Publications/Publication_Images/Figure6/Euler_Area46_depthsmooths.pdf", device = "pdf", dpi = 500, width = 1.5, height = 1.4)
```

**Anterior Cingulate (a24)**

```{r}
plot.depth.smooths(region = "L_a24_ROI", y_ticks = c(-0.01, 0, 0.01))
ggsave(filename = "/Users/valeriesydnor/Documents/ResearchProjects/CorticalMyelin_PlasticityDevelopment/Publications/Publication_Images/Figure6/Euler_Areaa24_depthsmooths.pdf", device = "pdf", dpi = 500, width = 1.5, height = 1.4)
```

## Principal component of developmental variability, based on trajectory curvatures


```{r}
#A function to compute the curvature of the estimated trajectory
curvature <- function(df, x,  y){
age.spline <- with(df, smooth.spline(x, y, df = 10))
first.deriv <- with(df, predict(age.spline, x = x, deriv = 1))
second.deriv <- with(df, predict(age.spline, x = x, deriv = 2))

k <- (second.deriv$y / ((1 + (first.deriv$y^2))^(3/2)))
k <- abs(k)
return(k)}
```

```{r}
#Calculate average curvature of R1 growth trajectories for every region at every cortical depth
trajectory.curvatures <- map_dfr(unique(gam.statistics.alldepths.long$orig_parcelname), function(r){
  region.smooths <- gam.smoothestimates.alldepths.long %>% filter(orig_parcelname == r) #smooth fits for this region at all depths
  depth.curvature <- region.smooths %>% group_by(depth) %>% do(curvature = mean(curvature(df = ., x = .$age , y = .$est))) %>% unnest(cols = c("curvature")) #calculate curvature of the smooth function
  depth.curvature$orig_parcelname <- r
  return(depth.curvature)
})

trajectory.curvatures <- merge(trajectory.curvatures, glasser.regions)
trajectory.curvatures.wide <- trajectory.curvatures %>% pivot_wider(id_cols = c("label", "orig_parcelname"), names_from = "depth", values_from = "curvature") 
```

```{r}
#run PCA
curvature.PCA <- trajectory.curvatures.wide %>% select(deep, middle, superficial) %>% prcomp(.)
```

```{r}
#variance explained
summary(curvature.PCA) 
```

```{r, warning = F, message = F}
curvature.PC1 <- curvature.PCA$x[,1] %>% as.data.frame %>% purrr::set_names("PC1")
curvature.PC1$PC1 <- as.numeric(scale(curvature.PC1$PC1))
curvature.PC1$label <- trajectory.curvatures.wide$label
curvature.PC1$orig_parcelname <- trajectory.curvatures.wide$orig_parcelname

ggplot() + 
   geom_brain(data = curvature.PC1, atlas = glasser, mapping = aes(fill = PC1), colour=I("gray10"), size=I(.08)) + theme_classic() +
 paletteer::scale_fill_paletteer_c("ggthemes::Purple", na.value = "white", direction = -1, limits = c(-1.25, .75), oob = squish) +  
  theme(legend.position = "none", axis.text = element_blank(), axis.line = element_blank(), axis.ticks = element_blank()) +
  new_scale_fill() + 
  geom_brain(data = glasser.plotting, atlas = glasser, mapping = aes(fill = cortex),  colour=I(alpha("gray50", 0)), size=I(0)) +
  scale_fill_manual(values = c(alpha("white", 0), "gray75")) 

ggsave(filename = "/Users/valeriesydnor/Documents/ResearchProjects/CorticalMyelin_PlasticityDevelopment/Publications/Publication_Images/Figure6_supplementary/Euler_PC1_curvature.pdf", device = "pdf", dpi = 500, width = 4.3, height = 2.2)
```

