---
title: "Frontal Layer Thickness in Histology and Ex-vivo MRI"
author: "Valerie Jill Sydnor"
output: 
  rmdformats::material:
    highlight: monochrome
    lightbox: true
    gallery: true
    css: /Volumes/Hera/Projects/corticalmyelin_development/code/corticalmyelin_maturation/results/custom_css/myelinanatomy.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(purrr)
library(cifti)
library(kableExtra)
```

```{r}
layers <- c("layer1", "layer2", "layer3", "layer4", "layer5", "layer6")
```

```{r}
glasser.regions <- read.csv("/Volumes/Hera/Projects/corticalmyelin_development/Maps/HCPMMP_glasseratlas/glasser360_regionlist.csv")
glasser.snr.exclude <- read.csv("/Volumes/Hera/Projects/corticalmyelin_development/Maps/SNR/glasser_SNR_exclusion.csv")
glasser.frontal <- read.csv("/Volumes/Hera/Projects/corticalmyelin_development/Maps/HCPMMP_glasseratlas/glasser360_regionlist_frontallobe.csv")
```

# Frontal Layer Thickness in BigBrain Histological Data (20 micrometer) 

Quantitative layer thickness from postmortern histology (BigBrain) obtained from  https://github.com/caseypaquola/BigBrainWarp, based on Wagstyl et al., 2020, Plos Biology 

```{r, warning = F}
layer1.cifti <- read_cifti("/Volumes/Hera/Projects/corticalmyelin_development/Maps/BigBrain_layerthickness/BBW_BigData/source-bigbrain_desc-layer1thickness_space-fsLR_den-32k.pscalar.nii")
layer2.cifti <- read_cifti("/Volumes/Hera/Projects/corticalmyelin_development/Maps/BigBrain_layerthickness/BBW_BigData/source-bigbrain_desc-layer2thickness_space-fsLR_den-32k.pscalar.nii")
layer3.cifti <- read_cifti("/Volumes/Hera/Projects/corticalmyelin_development/Maps/BigBrain_layerthickness/BBW_BigData/source-bigbrain_desc-layer3thickness_space-fsLR_den-32k.pscalar.nii")
layer4.cifti <- read_cifti("/Volumes/Hera/Projects/corticalmyelin_development/Maps/BigBrain_layerthickness/BBW_BigData/source-bigbrain_desc-layer4thickness_space-fsLR_den-32k.pscalar.nii")
layer5.cifti <- read_cifti("/Volumes/Hera/Projects/corticalmyelin_development/Maps/BigBrain_layerthickness/BBW_BigData/source-bigbrain_desc-layer5thickness_space-fsLR_den-32k.pscalar.nii")
layer6.cifti <- read_cifti("/Volumes/Hera/Projects/corticalmyelin_development/Maps/BigBrain_layerthickness/BBW_BigData/source-bigbrain_desc-layer6thickness_space-fsLR_den-32k.pscalar.nii")
```

```{r}
compute.layer.thickness <- function(layer){
  cifti.name <- sprintf("%s.cifti", layer)
  cifti <- get(cifti.name)
  layer.wholebrain <- data.frame(orig_parcelname = names(cifti$Parcel), thickness = cifti$data)
  layer.frontal <-  layer.wholebrain %>% filter(
                    orig_parcelname %in% glasser.frontal$orig_parcelname,
                    !orig_parcelname %in% glasser.snr.exclude$orig_parcelname)
  layer.frontal.meanthickness <- data.frame(layer = layer, thickness = mean(layer.frontal$thickness))
  return(layer.frontal.meanthickness)
}

laminar.thickness.histology <- map_dfr(layers, compute.layer.thickness)
```

```{r}
laminar.thickness.histology$thickness.proportion <- (laminar.thickness.histology$thickness / sum(laminar.thickness.histology$thickness))
```

```{r}
kable(laminar.thickness.histology)
```

**Layer 1 percent thickness**

```{r}
0
(laminar.thickness.histology$thickness.proportion[1])*100
```

**Layer 2/3 percent thickness**

```{r}
(laminar.thickness.histology$thickness.proportion[1])*100
(laminar.thickness.histology$thickness.proportion[1] + laminar.thickness.histology$thickness.proportion[2] + laminar.thickness.histology$thickness.proportion[3])*100
```

**Layer 4 percent thickness**

```{r}
(laminar.thickness.histology$thickness.proportion[1] + laminar.thickness.histology$thickness.proportion[2] + laminar.thickness.histology$thickness.proportion[3])*100
(laminar.thickness.histology$thickness.proportion[1] + laminar.thickness.histology$thickness.proportion[2] + laminar.thickness.histology$thickness.proportion[3] + laminar.thickness.histology$thickness.proportion[4])*100
```

**Layer 5/6 percent thickness**

```{r}
(laminar.thickness.histology$thickness.proportion[1] + laminar.thickness.histology$thickness.proportion[2] + laminar.thickness.histology$thickness.proportion[3] + laminar.thickness.histology$thickness.proportion[4])*100
(laminar.thickness.histology$thickness.proportion[1] + laminar.thickness.histology$thickness.proportion[2] + laminar.thickness.histology$thickness.proportion[3] + laminar.thickness.histology$thickness.proportion[4] + laminar.thickness.histology$thickness.proportion[5] + laminar.thickness.histology$thickness.proportion[6])*100
```

# Frontal Layer Thickness in Ex-vivo MRI (120 micrometer) 

Upper layers (1-3) proportion of thickness from ex-vivo MRI as quantified in Zeng et al., 2024, Cerebral Cortex

```{r}
upperlayer.thickness.exvivoMRI <- read.csv("/Volumes/Hera/Projects/corticalmyelin_development/Maps/ExvivoMRI_layerproportions/Frontal_supragranularlayer_percentages.csv")
upperlayer.thickness.exvivoMRI$hemisphere_average <- upperlayer.thickness.exvivoMRI %>% select(Left_hemisphere, Right_hemisphere) %>% rowMeans
```

```{r}
mean(upperlayer.thickness.exvivoMRI$hemisphere_average)
sd(upperlayer.thickness.exvivoMRI$hemisphere_average)
```


